---
title: "Analysis of Pressing using Tracking Data - Data Management"
author: "Pavanthi Sudasinghe"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Soccer Field

### Field Description

-   The width (along the x-axis) ranges from Xmin=−5250 to Xmax=5250, making the total width 5250−(−5250)=10,500 cm (105 m).

-   The height (along the y-axis) ranges from Ymin=−3400 to Ymax=3400, making the total height 3400−(−3400)=6800 cm (68 m).

-   The origin (0,0) represents the center of the field.

### Define the field dimensions

```{r dim}
x_min = -5250
x_max = 5250
y_min = -3400
y_max = 3400
```

### Plot the field

```{r plot}
# Plot the field boundary
plot(1, type = "n", xlim = c(x_min, x_max), ylim = c(y_min, y_max), 
     xlab = "X-coordinate", ylab = "Y-coordinate", 
     main = "Soccer Field", asp = 1)

# Draw the outer rectangle representing the field
rect(x_min, y_min, x_max, y_max, border = "darkgreen", lwd = 2) 

# Add the center line
lines(c(0, 0), c(y_min, y_max), col = "darkgreen", lwd = 2) 

# Add the center circle
center_circle_radius = 915   
symbols(0, 0, circles = center_circle_radius, inches = FALSE, 
        add = TRUE, lwd = 2, fg = "darkgreen")

# Draw penalty areas  
penalty_area_width = 4020
penalty_area_depth = 1650

# Left penalty area
rect(x_min, -penalty_area_width/2, x_min + penalty_area_depth, penalty_area_width/2, 
     border = "darkgreen", lty=2)

# Right penalty area
rect(x_max - penalty_area_depth, -penalty_area_width/2, x_max, penalty_area_width/2, 
     border = "darkgreen", lty=2)

# Add goal areas
goal_area_width = 1830
goal_area_depth = 550

# Left penalty area
rect(x_min, -goal_area_width/2, x_min + goal_area_depth, goal_area_width/2, 
     border = "darkgreen", lwd = 2)

# Right penalty area
rect(x_max - goal_area_depth, -goal_area_width/2, x_max, goal_area_width/2, 
     border = "darkgreen", lwd = 2)


# Add goal
goal_width = 732
goal_depth = 244

# Left goal
rect(x_min - goal_depth, -goal_width/2, x_min, goal_width/2, 
     border = "blue", lwd = 2)

# Right goal
rect(x_max, -goal_width/2, x_max + goal_depth, goal_width/2, 
     border = "blue", lwd = 2)

# Divide the field into custom zones
# Left
lines(c(-3000, -3000), c(-1500, 1500), col = "gray", lwd = 2, lty = 2)
lines(c(-5250, -3000), c(-1500, -1500), col = "gray", lwd = 2, lty = 2)
lines(c(-5250, -3000), c(1500, 1500), col = "gray", lwd = 2, lty = 2)

lines(c(0, -3000), c(-915, -3400), col = "gray", lwd = 2, lty = 2)
lines(c(-3000, -1800), c(-1500, -2400), col = "gray", lwd = 2, lty = 2)

lines(c(0, -3000), c(915, 3400), col = "gray", lwd = 2, lty = 2)
lines(c(-3000, -1800), c(1500, 2400), col = "gray", lwd = 2, lty = 2)

# Right
lines(c(3000, 3000), c(-1500, 1500), col = "gray", lwd = 2, lty = 2)
lines(c(5250, 3000), c(-1500, -1500), col = "gray", lwd = 2, lty = 2)
lines(c(5250, 3000), c(1500, 1500), col = "gray", lwd = 2, lty = 2)

lines(c(0, 3000), c(-915, -3400), col = "gray", lwd = 2, lty = 2)
lines(c(3000, 1800), c(-1500, -2400), col = "gray", lwd = 2, lty = 2)

lines(c(0, 3000), c(915, 3400), col = "gray", lwd = 2, lty = 2)
lines(c(3000, 1800), c(1500, 2400), col = "gray", lwd = 2, lty = 2)

# Define polygons and their labels
polygons = list(
  lb = list(x = c(-5250, -3000, -3000, -5250), y = c(-1500, -1500, 1500, 1500)),
  lc = list(x = c(-5250, -3000, -1800, -3000, -5250), y = c(1500, 1500, 2400, 3400, 3400)),
  la = list(x = c(-5250, -3000, -1800, -3000, -5250), y = c(-3400, -3400, -2400, -1500, -1500)),
  ld = list(x = c(-3000, 0, 0), y = c(-3400, -915, -3400)),
  lf = list(x = c(-3000, 0, 0), y = c(3400, 915, 3400)),
  le = list(x = c(-1800, 0, 0, -1800, -3000, -3000), y = c(-2400, -915, 915, 2400, 1500, -1500)),
  rb = list(x = c(5250, 3000, 3000, 5250), y = c(-1500, -1500, 1500, 1500)),
  rc = list(x = c(5250, 3000, 1800, 3000, 5250), y = c(1500, 1500, 2400, 3400, 3400)),
  ra = list(x = c(5250, 3000, 1800, 3000, 5250), y = c(-3400, -3400, -2400, -1500, -1500)),
  rd = list(x = c(3000, 0, 0), y = c(-3400, -915, -3400)),
  rf = list(x = c(3000, 0, 0), y = c(3400, 915, 3400)),
  re = list(x = c(1800, 0, 0, 1800, 3000, 3000), y = c(-2400, -915, 915, 2400, 1500, -1500))
)


for (name in names(polygons)) {
  # Draw the polygon
  polygon(polygons[[name]]$x, polygons[[name]]$y, border = "blue", col = adjustcolor("lightblue", alpha.f = 0.3))
  
  # Add labels for polygons
  text(mean(polygons[[name]]$x), mean(polygons[[name]]$y), labels = name, col = "darkblue", cex = 0.8)
  
  # # Add coordinates for each vertex
  # for (i in seq_along(polygons[[name]]$x)) {
  #   text(polygons[[name]]$x[i], polygons[[name]]$y[i], 
  #        labels = paste0("(", polygons[[name]]$x[i], ", ", polygons[[name]]$y[i], ")"), 
  #        col = "red", cex = 0.6, pos = 4)
  # }
}

```

Here, 'l' and 'r' represent 'left' and 'right', respectively.

### Importing Libraries

```{r libarries}
library(dplyr)
```

### Loading the tracking dataset

```{r datasets}
td = readRDS("C:/Users/pavan/Desktop/SFU/Research/Outputall/Outputall/2/tracking_df_2019-03-08-GZHOU-TTEDA.rds")
head(td)
```

### Loading the player dataset

```{r}
pd = read.csv("C:/Users/pavan/Desktop/SFU/Research/Outputall/Outputall/2/player_2019-03-08-GZHOU-TTEDA.csv",
                  header = TRUE)
head(pd)
```

### Loading the event dataset

```{r}
ed = read.csv(file_path <- "C:/Users/pavan/Desktop/SFU/Research/Outputall/Outputall/2/events_2019-03-08-GZHOU-TTEDA.csv",
                 header = TRUE)
head(ed)
```

### Cleaning tracking dataset

```{r}
# Remove unnecessary variables
td1 = subset(td, select = -c(.id, Z,NbPoints))

# Convert specific columns to numeric
td1[, c("T", "X", "Y", "IdActor", "IdHalf")] = lapply(td1[, c("T", "X", "Y", "IdActor", "IdHalf")], as.numeric)

# Filter players and ball data
td1_players = subset(td1 %>% filter(IsBall == 'False'), select = -c(IsBall))
td1_ball = subset(td1 %>% filter(IsBall == 'True'), select = -c(IsBall))

head(td1_players)
head(td1_ball)
```

### Calculating the speed of players

The analysis focuses on deriving the horizontal and vertical components of velocity and subsequently computing the overall speed.

A function named calculate_speed is defined to compute the speed of each player based on their positional data over time. The steps within the function are as follows:

-   Grouping by Player (IdActor): To ensure that speed is calculated for each player individually, the data is grouped by IdActor.

-   Velocity Components: Horizontal velocity (vx):

$$
\hat{v}_x(t_j) = \frac{x(t_j + \Delta) - x(t_j - \Delta)}{2\Delta}
$$

-   Vertical velocity (vy):

$$
\hat{v}_y(t_j) = \frac{y(t_j + \Delta) - y(t_j - \Delta)}{2\Delta}
$$

(<https://www.sfu.ca/~tswartz/papers/acceleration.pdf>)

-   Overall Speed (v): Derived as the Euclidean norm of the velocity components using the formula $$
    \hat{v} = \sqrt{\hat{v}_x^2 + \hat{v}_y^2}
    $$

```{r}
# Function to calculate speed 

# Function to determine delta based on T
get_delta = function(T) {
  if (T == 0) return(NA)  # Velocity should be NA at t=0
  else if (T == 100) return(100)
  else if (T == 200) return(200)
  else if (T == 300) return(300)
  else return(400)
}

# Function to compute velocity and speed based on variable delta
calculate_speed = function(data) {
  data %>%
    group_by(IdActor) %>%  # Group by player (IdActor)
    mutate(
      delta = sapply(T, get_delta),  # Get variable delta
      vx = (X[match(T + delta, T)] - X[match(T - delta, T)]) / (2 * delta),
      vy = (Y[match(T + delta, T)] - Y[match(T - delta, T)]) / (2 * delta),
      v = sqrt(vx^2 + vy^2)
    ) %>%
    ungroup() %>%
    select(-delta)  # Remove temporary column
}

# Apply the function to the datasets
td2_players = as.data.frame(calculate_speed(td1_players)) 

head(td2_players)
```

### Cleaning player dataset

```{r}
# keep necessary variables
pd_players = pd[, c('IdActor', 'IdTeam' ,'Name', 'Type', 'Position')]

# convert data type to numeric
pd_players$IdActor = as.numeric(pd_players$IdActor)
head(pd_players)
```

### Merging Player Data : Combining tracking player data and player data

```{r}
players_with_GK = left_join(td2_players, pd_players, by = "IdActor")
head(players_with_GK)
```

```{r}
# Removing rows where position is "goalkeeper"
players_without_GK = players_with_GK %>% filter(Position != "Goalkeeper")
head(players_without_GK)
```

### Assign geographic regions to each data point (player)

!!!This code takes around 1 hour to run

```{r}
# # Function to check if a point is inside the shape
# is_point_inside = function(px, py, x_coords, y_coords) {
#   n = length(x_coords)
#   inside = FALSE
#   j = n
#   for (i in 1:n) {
#     if (((y_coords[i] > py) != (y_coords[j] > py)) &&
#         (px < (x_coords[j] - x_coords[i]) * (py - y_coords[i]) / (y_coords[j] - y_coords[i]) + x_coords[i])) {
#       inside <- !inside
#     }
#     j = i
#   }
#   return(inside)
# }
# 
# players_without_GK$region = c()
# 
# # Assign regions
# for (i in 1:nrow(players_without_GK)) {
#   if (is.na(players_without_GK$X[i]) || is.na(players_without_GK$Y[i])) {
#     players_without_GK$region[i] <- NA
#   } else {
#     for (region in names(polygons)) {
#       poly <- polygons[[region]]
#       if (is_point_inside(players_without_GK$X[i], players_without_GK$Y[i], poly$x, poly$y)) {
#         players_without_GK$region[i] <- region
#         break
#       }
#     }
#   }
# }
# 
# 
# # Save the dataset
# saveRDS(players_without_GK, file = "players.Rds")
# ```
```


```{r}
players = readRDS("C:/Users/pavan/Desktop/SFU/Research/players.Rds")
head(players)
```


### Perform a merge on time (T) to align players and ball at the same time

```{r}
merged_player_ball = merge(players, td1_ball, by = c("T", "IdHalf"), suffixes = c("_player", "_ball"))
head(merged_player_ball) 
```

### Identifying Players in Possession of the Ball

To determine which players are in possession of the ball during the game, I assume that a player is near the ball if the euclidean distance between the player and the ball is less than a specified threshold (1.5m).

The steps are as follows:

1.  The distance between the player's position (X_player, Y_player) and the ball's position (X_ball, Y_ball) is calculated using this formula:

$$
\text{distance} = \sqrt{(X_{\text{player}} - X_{\text{ball}})^2 + (Y_{\text{player}} - Y_{\text{ball}})^2}.
$$

2.  Then retain only the players whose distance from the ball is less than 1.5 meters.

```{r}
radius = 150
near_ball_players = merged_player_ball %>%
  mutate(
    distance_ball_palyer = sqrt((X_player - X_ball)^2 + (Y_player - Y_ball)^2)
  ) %>%
  filter(distance_ball_palyer < radius)

head(near_ball_players)
```

```{r}
### Identify players in la, lc, ra, and rc
near_ball_players_ac = subset(near_ball_players, region %in% c("la", "lc", "ra", "rc"))
head(near_ball_players_ac)
```

### Identifying Players with Clear Space Around Them

To find players who have the ball and are in a clear space, I checked the distance between each player and all other players at the same time point. A player is considered to have clear space if all other players are more than 5 meters (500 cm) away from them.

The process is as follows:

1.  Grouping by Time: The data is grouped by the time points (T) to analyze player positions at each moment separately.

2.  Calculating Distances: For each player, the distance to all other players at the same time point is calculated.

3.  Checking Clear Space: A player is marked as having "clear space" if all distances to other players (except themselves) are greater than 5 meters.

4.  Filtering Results: Only players who have the ball and are in clear space are kept.

```{r}
# Define a function to calculate distance between two points
calculate_distance = function(x1, y1, x2, y2) {
  sqrt((x1 - x2)^2 + (y1 - y2)^2)
}

# Create a new column to check if the player has clear space
distance_players = near_ball_players_ac %>%
  group_by(T) %>%  # Group by time point
  mutate(
    clear_space = sapply(1:n(), function(i) {
      # Extract the coordinates of the current player
      current_x = X_player[i]
      current_y = Y_player[i]
      
      # Calculate distances to all other players at the same time point
      distances = calculate_distance(current_x, current_y, X_player, Y_player)
      
      # Check if all distances (excluding the player itself) are greater than 500cm (5m)
      all(distances[-i] > 500)  # -i to exclude the player itself
    })
  )

# Filter players with clear space
players_with_clear_space = as.data.frame(filter(distance_players, clear_space))
head(players_with_clear_space)
```

### Identify the team side

The x-coordinate of the home team's goalkeeper in the first half was identified from the dataset using filtering criteria based on player position, team type, and half identifier.

If the home team's goalkeeper's x-coordinate was negative, the home team was assumed to attack the right-side goal in the first half and the left-side goal in the second half, with the away team attacking the opposite goals.

```{r}
GKH1 = as.numeric(players_with_GK[players_with_GK$Position == 'Goalkeeper' & players_with_GK$Type == 'HomeTeam' & players_with_GK$IdHalf == '1', "X"][1])

# Assign team sides based on GH value
if (GKH1 < 0) {
  HomeTeam_1 = 'Right'; HomeTeam_2 = 'Left';   AwayTeam_1 = 'Left'; AwayTeam_2 = 'Right'
} else {
  HomeTeam_1 = 'Left'; HomeTeam_2 = 'Right'; AwayTeam_1 = 'Right'; AwayTeam_2 = 'Left'
}

# Print the result
cat("Home Team - First Half:", HomeTeam_1, "\n",
    "Away Team - First Half:", AwayTeam_1, "\n",
    "Home Team - Second Half:", HomeTeam_2, "\n",
    "Away Team - Second Half:", AwayTeam_2, "\n")
```

### Filtering Player Positions by Goal Orientation and Match Half

This analysis examines the distribution of players in specific regions of the field, classified by team type (Home or Away) and match half (First or Second).

```{r}
if (GKH1<0){
  r_1_H = players_with_clear_space[players_with_clear_space$region %in% c('ra', 'rc') & players_with_clear_space$IdHalf == 1 & players_with_clear_space$Type == 'HomeTeam', ]
  r_1_A = players[players$region %in% c('ra', 'rc', 'rb', 'rd', 're', 'rf') & players$IdHalf == 1 & players$Type == 'AwayTeam', ]
  
  r_2_A = players_with_clear_space[players_with_clear_space$region %in% c('ra', 'rc') & players_with_clear_space$IdHalf == 2 & players_with_clear_space$Type == 'AwayTeam', ]
  r_2_H = players[players$region %in% c('ra', 'rc', 'rb', 'rd', 're', 'rf') & players$IdHalf == 2 & players$Type == 'HomeTeam', ]
  
  l_1_A = players_with_clear_space[players_with_clear_space$region %in% c('la', 'lc') & players_with_clear_space$IdHalf == 1 & players_with_clear_space$Type == 'AwayTeam', ]
  l_1_H = players[players$region %in% c('la', 'lc', 'lb', 'ld', 'le', 'lf') & players$IdHalf == 1 & players$Type == 'HomeTeam', ]
  
  l_2_H = players_with_clear_space[players_with_clear_space$region %in% c('la', 'lc') & players_with_clear_space$IdHalf == 2 & players_with_clear_space$Type == 'HomeTeam', ]
  l_2_A = players[players$region %in% c('la', 'lc', 'lb', 'ld', 'le', 'lf') & players$IdHalf == 2 & players$Type == 'AwayTeam', ]
  
}else{
  
  r_1_A = players_with_clear_space[players_with_clear_space$region %in% c('ra', 'rc') & players_with_clear_space$IdHalf == 1 & players_with_clear_space$Type ==
'AwayTeam', ]
  r_1_H = players[players$region %in% c('ra', 'rc', 'rb', 'rd', 're', 'rf') & players$IdHalf == 1 & players$Type == 'HomeTeam', ]
  
  r_2_H = players_with_clear_space[players_with_clear_space$region %in% c('ra', 'rc') & players_with_clear_space$IdHalf == 2 & players_with_clear_space$Type ==
'HomeTeam', ]
  r_2_A = players[players$region %in% c('ra', 'rc', 'rb', 'rd', 're', 'rf') & players$IdHalf == 2 & players$Type == 'AwayTeam', ]
  
  l_1_H = players_with_clear_space[players_with_clear_space$region %in% c('la', 'lc') & players_with_clear_space$IdHalf == 1 & players_with_clear_space$Type ==
'HomeTeam', ]
  l_1_A = players[players$region %in% c('la', 'lc', 'lb', 'ld', 'le', 'lf') & players$IdHalf == 1 & players$Type == 'AwayTeam', ]
  
  l_2_A = players_with_clear_space[players_with_clear_space$region %in% c('la', 'lc') & players_with_clear_space$IdHalf == 2 & players_with_clear_space$Type ==
'AwayTeam', ]
  l_2_H = players[players$region %in% c('la', 'lc', 'lb', 'ld', 'le', 'lf') & players$IdHalf == 2 & players$Type == 'HomeTeam', ]
  
}


```

### Defining the Speed Threshold for a Presser

To identify pressers (from the opposing team) moving towards a player with the ball at a speed greater than 30 km/hr, I converted the speed into the same units as the dataset, which uses centimeters (cm) for position and milliseconds (ms) for time.

```{r}
# Given speed in km/hr
speed_km_hr = 30

# Convert km/hr to cm/ms
speed_cm_ms = ((speed_km_hr* 1000 / 3600)/1000) * 100

# Output
speed_threshold = speed_cm_ms
speed_threshold
```

### Filtering Speed Data Based on a Threshold

The following code is used to filter the computed speed data to retain only the rows where the overall speed (v) exceeds a predefined threshold (30 Km/hr):

```{r}
if(GKH1<0){
  filtered_r1A = r_1_A[r_1_A$v > speed_threshold,] 
  filtered_l1H = l_1_H[l_1_H$v > speed_threshold,]  
  filtered_r2H = r_2_H[r_2_H$v > speed_threshold,]  
  filtered_l2A = l_2_A[l_2_A$v > speed_threshold,] 
}else{
  filtered_l1A = l_1_A[l_1_A$v > speed_threshold,] 
  filtered_r1H = r_1_H[r_1_H$v > speed_threshold,]  
  filtered_l2H = l_2_H[l_2_H$v > speed_threshold,]  
  filtered_r2A = r_2_A[r_2_A$v > speed_threshold,]   
}
```

### Merging Speed Data of Pressing Players with Ball Possession Data

The following code merges two datasets to align the speed of pressing team players with the positional data of players who have possession of the ball:

```{r}
if(GKH1<0){
  merged_datar1 = merge(filtered_r1A, r_1_H, by = "T" ) 
  merged_datal1 = merge(filtered_l1H, l_1_A, by = "T" ) 
  merged_datar2 = merge(filtered_r2H, r_2_A, by = "T" ) 
  merged_datal2 = merge(filtered_l2A, l_2_H, by = "T" ) 
}else{
  merged_datal1 = merge(filtered_l1A, l_1_H, by = "T" ) 
  merged_datar1 = merge(filtered_r1H, r_1_A, by = "T" ) 
  merged_datal2 = merge(filtered_l2H, l_2_A, by = "T" ) 
  merged_datar2 = merge(filtered_r2A, r_2_H, by = "T" )  
}

# Combine all datasets into one
data = rbind(merged_datar1, merged_datal1, merged_datar2, merged_datal2)
head(data)
```

### Function for calculating the angle between two vectors representing pressing player and the ball carrier

<https://www.nature.com/articles/s41598-024-62480-7>

Compute the Direction Vector (d)

The direction vector (dx,dy) represents the displacement from the pressing player's current position (X,Y) to the ball carrier's position (X player,Y player ):

$$
dx = X_{\text{player}}-X
$$

$$
dy = Y_{\text{player}} - Y
$$

Extract the Player's Velocity Vector ( v)

The velocity components (vx,vy) of the player are obtained directly from the dataset.

Calculate the Dot Product

The dot product of the direction vector and velocity vector is computed as follows:

\$\$

uv = dx.vx+dy.vy

\$\$ Compute the Magnitudes

The magnitudes of both vectors are calculated using the Euclidean norm:

\$\$

u = \sqrt{dx^2 + dy^2}

v = \sqrt{vx^2 + vy^2}

\$\$ Determine the Angle

The angle θ between the two vectors is derived using the dot product formula: $$
\theta = \arccos\left( \frac{u \cdot v}{|u| |v|} \right)
$$

To prevent division by zero, cases where u⋅v=0 are assigned NA values.

The result is expressed in degrees for easier interpretation. If the angle is less than or equal to 90 degrees, it implies that the player is running towards the target, otherwise, the player is running away.

```{r}
calculate_direction = function(data, x_col, y_col, vx_col, vy_col, target_x_col, target_y_col) {
  if (nrow(data) == 0) return(data)  # Handle empty dataframe case
  
  dx = data[[target_x_col]] - data[[x_col]]
  dy = data[[target_y_col]] - data[[y_col]]
  vx = data[[vx_col]]
  vy = data[[vy_col]]

  uv = dx * vx + dy * vy
  u = sqrt(dx^2 + dy^2)
  v = sqrt(vx^2 + vy^2)

  data$angle = rep(NA, nrow(data))  # Initialize
  valid_indices = which(u * v != 0) # Avoid division by zero
  cos_theta = pmax(pmin(uv[valid_indices] / (u[valid_indices] * v[valid_indices]), 1), -1)
  data$angle[valid_indices] = acos(cos_theta) * (180 / pi)

  data$Running_towards = ifelse(data$angle <= 90, 'Yes', 'No')
  
  return(data)
}


# Apply the function to each merged dataset
Final = calculate_direction(data, "X", "Y", "vx", "vy", "X_player", "Y_player")
head(Final)
nrow(Final)
```

## Defining the variables required for the final dataset

1.  match id

```{r}
Final$match_id = rep(tools::file_path_sans_ext(substr(basename(file_path), 8, nchar(basename(file_path)))),nrow(Final))
```

2.  time of the match

```{r}
Final$time = Final$T
```

3.  team id corresponding to the home team

```{r}
Final$Home_Team = rep(pd$IdTeam[pd$Type=='HomeTeam'][1], nrow(Final))
```

4.  team1 id corresponding to the pressed team (refers to the team that is under pressure from the opponent)

```{r}
# Create the table as a data frame
table_data <- data.frame(
  "GKH1 < 0" = c("Right (ra, rc)", "Left (la, lc)", "Left (la, lc)", "Right (ra, rc)"),
  "GKH1 >= 0" = c("Left (la, lc)", "Right (ra, rc)", "Right (ra, rc)", "Left (la, lc)")
)

# Assign row names
rownames(table_data) <- c("First Half: Home Team Pressed", 
                          "First Half: Away Team Pressed", 
                          "Second Half: Home Team Pressed", 
                          "Second Half: Away Team Pressed")

# Print the table
print(table_data)
```

```{r}
# Initialize team1_id with 0
Final$team1 = 0

# Update team1_id based on the conditions

if(GKH1<0){
  Final$team1 = ifelse(
  Final$IdHalf.y == 1 & Final$Type.y == 'HomeTeam' & Final$region.y %in% c('ra', 'rc'),
  pd$IdTeam[pd$Type == 'HomeTeam'][1],
  ifelse(
    Final$IdHalf.y == 1 & Final$Type.y == 'AwayTeam' & Final$region.y %in% c('la', 'lc'),
    pd$IdTeam[pd$Type == 'AwayTeam'][1],
    ifelse(
      Final$IdHalf.y == 2 & Final$Type.y == 'AwayTeam' & Final$region.y %in% c('ra', 'rc'),
      pd$IdTeam[pd$Type == 'AwayTeam'][1],
      ifelse(
        Final$IdHalf.y == 2 & Final$Type.y == 'HomeTeam' & Final$region.y %in% c('la', 'lc'),
        pd$IdTeam[pd$Type == 'HomeTeam'][1],
        0 # Default to 0 if no conditions are met
      )
    )
  )
)
}else{
  Final$team1 = ifelse(
  Final$IdHalf.y == 1 & Final$Type.y == 'HomeTeam' & Final$region.y %in% c('la', 'lc'),
  pd$IdTeam[pd$Type == 'HomeTeam'][1],
  ifelse(
    Final$IdHalf.y == 1 & Final$Type.y == 'AwayTeam' & Final$region.y %in% c('ra', 'rc'),
    pd$IdTeam[pd$Type == 'AwayTeam'][1],
    ifelse(
      Final$IdHalf.y == 2 & Final$Type.y == 'AwayTeam' & Final$region.y %in% c('la', 'lc'),
      pd$IdTeam[pd$Type == 'AwayTeam'][1],
      ifelse(
        Final$IdHalf.y == 2 & Final$Type.y == 'HomeTeam' & Final$region.y %in% c('ra', 'rc'),
        pd$IdTeam[pd$Type == 'HomeTeam'][1],
        0 # Default to 0 if no conditions are met
      )
    )
  )
)
}



```

5.  team2 id corresponding to the pressing team (refers to the team that is applying pressure on the opponent)

```{r}
# Create the table as a data frame
table_data <- data.frame(
  "GKH1 < 0" = c("Left", "Right",  "Right", "Left"),
  "GKH1 >= 0" = c("Right", "Left", "Left" ,  "Right")
)

# Assign row names
rownames(table_data) <- c("First Half: Home Team Pressing", 
                          "First Half: Away Team Pressing", 
                          "Second Half: Home Team Pressing", 
                          "Second Half: Away Team Pressing")

# Print the table
print(table_data)
```

```{r}
# Initialize team2_id with 0
Final$team2 = 0

if(GKH1<0){
  # Update team2_id based on the conditions
Final$team2_id = ifelse(
  Final$IdHalf.x == 1 & Final$Type.x == 'HomeTeam' & Final$region.x %in% c('la', 'lb', 'lc', 'ld', 'le', 'lf'),
  pd$IdTeam[pd$Type == 'HomeTeam'][1],
  ifelse(
    Final$IdHalf.x == 1 & Final$Type.x == 'AwayTeam' & Final$region.x %in% c('ra', 'rb', 'rc', 'rd', 're', 'rf'),
    pd$IdTeam[pd$Type == 'AwayTeam'][1],
    ifelse(
      Final$IdHalf.x == 2 & Final$Type.x == 'AwayTeam' & Final$region.x %in% c('la', 'lb', 'lc', 'ld', 'le', 'lf'),
      pd$IdTeam[pd$Type == 'AwayTeam'][1],
      ifelse(
        Final$IdHalf.x == 2 & Final$Type.x == 'HomeTeam' & Final$region.x %in% c('ra', 'rb', 'rc', 'rd', 're', 'rf'),
        pd$IdTeam[pd$Type == 'HomeTeam'][1],
        0 # Default to 0 if no conditions are met
      )
    )
  )
)
}else{
  # Update team2_id based on the conditions
Final$team2 = ifelse(
  Final$IdHalf.x == 1 & Final$Type.x == 'HomeTeam' & Final$region.x %in% c('ra', 'rb', 'rc', 'rd', 're', 'rf'),
  pd$IdTeam[pd$Type == 'HomeTeam'][1],
  ifelse(
    Final$IdHalf.x == 1 & Final$Type.x == 'AwayTeam' & Final$region.x %in% c('la', 'lb', 'lc', 'ld', 'le', 'lf'),
    pd$IdTeam[pd$Type == 'AwayTeam'][1],
    ifelse(
      Final$IdHalf.x == 2 & Final$Type.x == 'AwayTeam' & Final$region.x %in% c('ra', 'rb', 'rc', 'rd', 're', 'rf'),
      pd$IdTeam[pd$Type == 'AwayTeam'][1],
      ifelse(
        Final$IdHalf.x == 2 & Final$Type.x == 'HomeTeam' & Final$region.x %in% c('la', 'lb', 'lc', 'ld', 'le', 'lf'),
        pd$IdTeam[pd$Type == 'HomeTeam'][1],
        0 # Default to 0 if no conditions are met
      )
    )
  )
)
}

```

6.  Player ID corresponding to the player who is pressed in region a or c (refers to the player who is under pressure from an opponent)

```{r}
Final$player_id1 = Final$IdActor_player
```

7.  Player ID corresponding to the player who pressed (refers to the player who is applying pressure on an opponent)

```{r}
Final$player_id2 = Final$IdActor
```

8.  goals scored for team1 at time of press

-   If the event occurs in the first half and the Home Team is being pressed, the code looks for the Home Team’s score in the ed data frame. Specifically, it finds the score corresponding to the earliest time that occurs after the pressing event (Final\$T[i]) during the first half. If a matching score is found, it updates the goals_t1 column for that row.

-   If the event occurs in the first half and the Away Team is being pressed, the procedure is the same as for the Home Team, except this time it retrieves the Away Team’s score from the ed data frame.

-   The same logic applies for the Home Team and Away Team during the second half.

```{r}
# Initialize the column
Final$goals_t1 = 0

# Iterate through each row of the Final data frame
for (i in 1:nrow(Final)) {
  
  # First Half and Home Team
  if (Final$IdHalf.y[i] == 1 && Final$Type.y[i] == 'HomeTeam') {
    match = ed$ScoreHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >= 1) {
      Final$goals_t1[i] = match[1]
    } else {
      Final$goals_t1[i] = NA
    }
  
  # First Half and Away Team
  } else if (Final$IdHalf.y[i] == 1 && Final$Type.y[i] == 'AwayTeam') {
    match = ed$ScoreAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >=1 ) {
      Final$goals_t1[i] = match[1]
    } else {
      Final$goals_t1[i] = NA
    }
  
  # Second Half and Away Team
  } else if (Final$IdHalf.y[i] == 2 && Final$Type.y[i] == 'AwayTeam') {
    match = ed$ScoreAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$goals_t1[i] = match[1]
    } else {
      Final$goals_t1[i] = NA
    }
  
  # Second Half and Home Team
  } else if (Final$IdHalf.y[i] == 2 && Final$Type.y[i] == 'HomeTeam') {
    match = ed$ScoreHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$goals_t1[i] = match[1]
    } else {
      Final$goals_t1[i] = NA
    }
  
  # Default case (in case no conditions are met)
  } else {
    Final$goals_t1[i] = NA
  }
}
```

9.  goals scored for team2 at time of press

Similar to the previous step, goals scored by Team 2 at the time of pressing are extracted.

```{r}
# Initialize goals_t2 with 0
Final$goals_t2 = 0

# Iterate through each row of the Final data frame
for (i in 1:nrow(Final)) {
  
  # First Half and Home Team
  if (Final$IdHalf.x[i] == 1 && Final$Type.x[i] == 'HomeTeam') {
    match = ed$ScoreHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >= 1) {
      Final$goals_t2[i] = match[1]
    } else {
      Final$goals_t2[i] = NA
    }
  
  # First Half and Away Team
  } else if (Final$IdHalf.x[i] == 1 && Final$Type.x[i] == 'AwayTeam') {
    match = ed$ScoreAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >= 1) {
      Final$goals_t2[i] = match[1]
    } else {
      Final$goals_t2[i] = NA
    }
  
  # Second Half and Away Team
  } else if (Final$IdHalf.x[i] == 2 && Final$Type.x[i] == 'AwayTeam') {
    match = ed$ScoreAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$goals_t2[i] = match[1]
    } else {
      Final$goals_t2[i] = NA
    }
  
  # Second Half and Home Team
  } else if (Final$IdHalf.x[i] == 2 && Final$Type.x[i] == 'HomeTeam') {
    match = ed$ScoreHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$goals_t2[i] = match[1]
    } else {
      Final$goals_t2[i] = NA
    }
  
  # Default case (in case no conditions are met)
  } else {
    Final$goals_t2[i] = NA
  }
}
```

10. ID half

```{r}
Final$half = Final$IdHalf.x
```

11. red_cards_t1 corresponding to the number of red cards Team 1 had at the time of the press. (Refers to the number of red cards the team under pressure had at the time of the press.)

```{r}
# Initialize the column
Final$red_cards_t1 = 0

# Iterate through each row of the Final data frame
for (i in 1:nrow(Final)) {
  
  # First Half and Home Team
  if (Final$IdHalf.y[i] == 1 && Final$Type.y[i] == 'HomeTeam') {
    match = ed$RedCardsHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >= 1) {
      Final$red_cards_t1[i] = match[1]
    } else {
      Final$red_cards_t1[i] = NA
    }
  
  # First Half and Away Team
  } else if (Final$IdHalf.y[i] == 1 && Final$Type.y[i] == 'AwayTeam') {
    match = ed$RedCardsAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >=1 ) {
      Final$red_cards_t1[i] = match[1]
    } else {
      Final$red_cards_t1[i] = NA
    }
  
  # Second Half and Away Team
  } else if (Final$IdHalf.y[i] == 2 && Final$Type.y[i] == 'AwayTeam') {
    match = ed$RedCardsAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$red_cards_t1[i] = match[1]
    } else {
      Final$red_cards_t1[i] = NA
    }
  
  # Second Half and Home Team
  } else if (Final$IdHalf.y[i] == 2 && Final$Type.y[i] == 'HomeTeam') {
    match = ed$RedCardsHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$red_cards_t1[i] = match[1]
    } else {
      Final$red_cards_t1[i] = NA
    }
  
  # Default case (in case no conditions are met)
  } else {
    Final$red_cards_t1[i] = NA
  }
}
```

12. red_cards_t2 corresponding to the number of red cards Team 2 had at the time of the press. (Refers to the number of red cards the team applying pressure had at the time of the press.)

```{r}
# Initialize red_cards_t2 with 0
Final$red_cards_t2 = 0

# Iterate through each row of the Final data frame
for (i in 1:nrow(Final)) {
  
  # First Half and Home Team
  if (Final$IdHalf.x[i] == 1 && Final$Type.x[i] == 'HomeTeam') {
    match = ed$RedCardsHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >= 1) {
      Final$red_cards_t2[i] = match[1]
    } else {
      Final$red_cards_t2[i] = NA
    }
  
  # First Half and Away Team
  } else if (Final$IdHalf.x[i] == 1 && Final$Type.x[i] == 'AwayTeam') {
    match = ed$RedCardsAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 1])]
    if (length(match) >= 1) {
      Final$red_cards_t2[i] = match[1]
    } else {
      Final$red_cards_t2[i] = NA
    }
  
  # Second Half and Away Team
  } else if (Final$IdHalf.x[i] == 2 && Final$Type.x[i] == 'AwayTeam') {
    match = ed$RedCardsAwayTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$red_cards_t2[i] = match[1]
    } else {
      Final$red_cards_t2[i] = NA
    }
  
  # Second Half and Home Team
  } else if (Final$IdHalf.x[i] == 2 && Final$Type.x[i] == 'HomeTeam') {
    match = ed$RedCardsHomeTeam[ed$Time == min(ed$Time[ed$Time > Final$T[i] & ed$IdHalf == 2])]
    if (length(match) >= 1) {
      Final$red_cards_t2[i] = match[1]
    } else {
      Final$red_cards_t2[i] = NA
    }
  
  # Default case (in case no conditions are met)
  } else {
    Final$red_cards_t2[i] = NA
  }
}
```



13. 

```{r}
# ed_new = ed[,c("Time", "IdHalf", "IdActor1", "EventName", "BodyPart", "LocationX", "LocationY")]
# 
# 
# # Function to check if a point is inside the shape
# is_point_inside = function(px, py, x_coords, y_coords) {
#   n = length(x_coords)
#   inside = FALSE
#   j = n
#   for (i in 1:n) {
#     if (((y_coords[i] > py) != (y_coords[j] > py)) &&
#         (px < (x_coords[j] - x_coords[i]) * (py - y_coords[i]) / (y_coords[j] - y_coords[i]) + x_coords[i])) {
#       inside <- !inside
#     }
#     j = i
#   }
#   return(inside)
# }
# 
# ed_new$region = c()
# 
# # Assign regions
# for (i in 1:nrow(ed_new)) {
#   if (is.na(ed_new$LocationX[i]) || is.na(ed_new$LocationY[i])) {
#     ed_new$region[i] <- NA
#   } else {
#     for (region in names(polygons)) {
#       poly <- polygons[[region]]
#       if (is_point_inside(ed_new$LocationX[i], ed_new$LocationY[i], poly$x, poly$y)) {
#         ed_new$region[i] <- region
#         break
#       }
#     }
#   }
# }
# 
# 
# # Save the dataset
# saveRDS(ed_new, file = "ed_new_events.Rds")

ed_events = readRDS("C:/Users/pavan/Desktop/SFU/Research/ed_new_events.Rds")
Final <- Final[order( Final$half, Final$time), ]
ed_events_sorted <- ed_events[order( ed$IdHalf, ed$Time), ]
# Create empty columns for event names and regions
Final$event <- NA 

# Loop over each row in Final_dataset_sorted
for (i in seq_len(nrow(Final))) {
  
  # Find first time in ed_events_sorted greater than Final_dataset_sorted$time[i]
  matched_times_ed <- ed_events_sorted$Time[ed_events_sorted$Time > Final$time[i]]
  
  if (length(matched_times_ed) > 0) {
    matched_time_ed <- matched_times_ed[1]  # Take the first match
    
    # Find rows in Final_dataset_sorted that have time < matched_time_ed and same half
    matched_rows <- which(Final$time < matched_time_ed & 
                           Final$half == ed_events_sorted$IdHalf[i])  # Use ed_events_sorted$IdHalf[i] for current event
    
    if (length(matched_rows) > 0) {
      matched_time_F = max(Final$time[matched_rows], na.rm = TRUE)
      matched_row_F <- which(Final$time == matched_time_F)
      
      # Find row index in ed_events_sorted that matches matched_time_ed
      matched_row_ed <- which(ed_events_sorted$Time == matched_time_ed)
      
      if (length(matched_row_ed) > 0) {
        matched_event <- ed_events_sorted$EventName[matched_row_ed]
         
        
        # Assign values to the corresponding row in Final_dataset_sorted
        Final$event[matched_row_F] <- matched_event
         
      }
    }
  }
}

# Print the updated dataframe
head(Final)


```





### Assign geographic regions to each data point (ball)

!!!This code takes around 1 hour to run

```{r}
# Function to check if a point is inside the shape
is_point_inside = function(px, py, x_coords, y_coords) {
  n = length(x_coords)
  inside = FALSE
  j = n
  for (i in 1:n) {
    if (((y_coords[i] > py) != (y_coords[j] > py)) &&
        (px < (x_coords[j] - x_coords[i]) * (py - y_coords[i]) / (y_coords[j] - y_coords[i]) + x_coords[i])) {
      inside <- !inside
    }
    j = i
  }
  return(inside)
}

td1_ball$region = c()

# Assign regions
for (i in 1:nrow(td1_ball)) {
  if (is.na(td1_ball$X[i]) || is.na(td1_ball$Y[i])) {
    td1_ball$region[i] <- NA
  } else {
    for (region in names(polygons)) {
      poly <- polygons[[region]]
      if (is_point_inside(td1_ball$X[i], td1_ball$Y[i], poly$x, poly$y)) {
        td1_ball$region[i] <- region
        break
      }
    }
  }
}


# Save the dataset
saveRDS(td1_ball, file = "ball.Rds")
```


```{r}
ball = readRDS("C:/Users/pavan/Desktop/SFU/Research/ball.Rds")
head(ball)
```
```{r}
Final$IdHalf = Final$IdHalf.x
Final_merged = merge(Final, ball, by = c("IdHalf", "T")) 
head(Final_merged)
```




## Final dataset (up to this point)

Keep only the necessary variables.

```{r}
Final_dataset = Final_merged[, c('match_id', 'time', 'half','Home_Team', 'team1', 'team2', 'player_id1', 'player_id2',  'goals_t1', 'goals_t2', 'red_cards_t1', 'red_cards_t2', 'event', 'region')] 

Final_dataset_sorted = Final_dataset[order( Final_dataset$half, Final_dataset$time), ]

head(Final_dataset_sorted)

```
